---
title: "Project Work"
---

Research Question: What is the relationship between acoustic features and popularity of a song? Are their trends in features that increase the popularity of a song?

**Variables needed for characteristics / acoustic features:**
(audio_features)
acousticness
danceability
energy
instrumentalness
key
liveness
loudness
speechiness
spotify_track_duration_ms
tempo
time_signature
valence
song


**Variables needed for popularity measurement:**
(audio_features)
spotify_track_popularity


Filter out songs that do not have audio feature data. 

```{r}
#Imports
library(tidyverse)
```
```{r}
#read csv
audioFeatures <- read.csv("data/audio_features.csv")
```

```{r}
#filter to get songs with a popularity score > 85
audioFeaturesTop <- audioFeatures |>
  filter(spotify_track_popularity > 85)
#filter to get songs with a popularity score < 1
audioFeaturesBottom <- audioFeatures |>
  filter(spotify_track_popularity < 1)
```


```{r}
summary(audioFeaturesTop)
```
```{r}
summary(audioFeaturesBottom)
```

























