---
title: "Project Work"
---

Research Question: What is the relationship between acoustic features and popularity of a song? Are their trends in features that increase the popularity of a song?

**Variables needed for characteristics / acoustic features:**
(audio_features)
acousticness
danceability
energy
instrumentalness
key
liveness
loudness
speechiness
spotify_track_duration_ms
tempo
time_signature
valence
song


**Variables needed for popularity measurement:**
(audio_features)
spotify_track_popularity


Filter out songs that do not have audio feature data. 

```{r}
#Imports
library(tidyverse)
library(viridis)
```
```{r}
#read csv
audioFeatures <- read.csv("data/audio_features.csv")
```


```{r}
audioFeatures <- audioFeatures |>
  mutate(max = pmax(danceability,energy,valence,liveness,instrumentalness,acousticness,speechiness)) 
  
#removed duplicate entries
audioFeatures = audioFeatures[!duplicated(audioFeatures$song),]
  
#adds max and maxCat variables
audioFeatures <- audioFeatures |>
  mutate(maxCat = ifelse(danceability == max,"danceability",
                   ifelse(energy == max,"energy",
                   ifelse(valence == max,"valence",
                   ifelse(liveness == max,"liveness",
                   ifelse(instrumentalness == max,"instrumentalness",
                   ifelse(acousticness == max,"acousticness",
                   ifelse(speechiness == max,"speechiness",
                          "N/A"))))))))


#filter to get songs with a popularity score > 85
audioFeaturesTop <- audioFeatures |>
  filter(spotify_track_popularity > 85)
#filter to get songs with a popularity score < 15
audioFeaturesBottom <- audioFeatures |>
  filter(spotify_track_popularity < 15 )
```


```{r}
summary(audioFeaturesTop)
```
```{r}
summary(audioFeaturesBottom)
```

```{r}
audioFeaturesTop %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point() + 
  labs(title = "Top Songs")+
  ylab("Spotify Track Popultaity")+
  xlab("Max")+
  theme(plot.title = element_text(size=22, face = "bold"), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        legend.title = element_text(face = "bold"),
        plot.background = element_rect(fill = "gray")
        )

```
```{r}
audioFeaturesBottom %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point() + 
  labs(title = "Botttom Songs")+
  ylab("Spotify Track Popultaity")+
  xlab("Max")+
  theme(plot.title = element_text(size=22, face = "bold"), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        legend.title = element_text(face = "bold"),
        plot.background = element_rect(fill = "gray")
        )

```
```{r}
audioFeatures %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point()
```

```{r}
audioFeatures |>
  filter(spotify_track_popularity == 100) |>
  filter(song == "Dance Monkey") |>
  select(song,danceability,energy,valence,liveness,instrumentalness,acousticness,speechiness) |>
  pivot_longer(cols = -song, names_to = "Characteristic", values_to = "Values") |>
  
  ggplot(aes(x = Characteristic, y = Values, fill = Characteristic)) +
  geom_col() +
  ylim(c(0, 1)) +
  scale_fill_manual(values = c("#ccff33", "#9ef01a", "#9ef01a", "#38b000", "#008000", "#007200", "#006400")) +
  labs(title = "Dance Monkey Audio Features", subtitle = "Top Song") +
  theme(plot.title = element_text(size=22, face = "bold"), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        legend.title = element_text(face = "bold"),
        plot.background = element_rect(fill = "gray")
        )
```
```{r}
audioFeatures |>
  filter(spotify_track_popularity == 0) |>
  filter(song == "Goldfinger") |>
  select(song,danceability,energy,valence,liveness,instrumentalness,acousticness,speechiness) |>
  pivot_longer(cols = -song, names_to = "Characteristic", values_to = "Values") |>
  
  ggplot(aes(x = Characteristic, y = Values, fill = Characteristic)) +
  geom_col() +
  ylim(c(0, 1)) +
  scale_fill_manual(values = c("#ccff33", "#9ef01a", "#9ef01a", "#38b000", "#008000", "#007200", "#006400")) +
  labs(title = "Goldfinger Audio Features", subtitle = "Bottom Song") +
  theme(plot.title = element_text(size=22, face = "bold"), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        legend.title = element_text(face = "bold"),
        plot.background = element_rect(fill = "gray")
        )
```



















