---
title: "Project Work"
---

Research Question: What is the relationship between acoustic features and popularity of a song? Are their trends in features that increase the popularity of a song?

**Variables needed for characteristics / acoustic features:**
(audio_features)
acousticness
danceability
energy
instrumentalness
key
liveness
loudness
speechiness
spotify_track_duration_ms
tempo
time_signature
valence
song


**Variables needed for popularity measurement:**
(audio_features)
spotify_track_popularity


Filter out songs that do not have audio feature data. 

```{r}
#Imports
library(tidyverse)
```
```{r}
#read csv
audioFeatures <- read.csv("data/audio_features.csv")
```


```{r}
audioFeatures <- audioFeatures |>
  mutate(max = pmax(danceability,energy,valence,liveness,instrumentalness,acousticness,speechiness))

audioFeatures <- audioFeatures |>
  mutate(maxCat = ifelse(danceability == max,"danceability",
                   ifelse(energy == max,"energy",
                   ifelse(valence == max,"valence",
                   ifelse(liveness == max,"liveness",
                   ifelse(instrumentalness == max,"instrumentalness",
                   ifelse(acousticness == max,"acousticness",
                   ifelse(speechiness == max,"speechiness",
                          "N/A"))))))))
```


```{r}
#filter to get songs with a popularity score > 85
audioFeaturesTop <- audioFeatures |>
  filter(spotify_track_popularity > 85)
#filter to get songs with a popularity score < 1
audioFeaturesBottom <- audioFeatures |>
  filter(spotify_track_popularity < 1)
```


```{r}
summary(audioFeaturesTop)
```
```{r}
summary(audioFeaturesBottom)
```

```{r}
audioFeaturesTop %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point()

```
```{r}
audioFeaturesBottom %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point()

```
```{r}
audioFeatures %>%
  ggplot(aes(x = max, y = spotify_track_popularity, color = maxCat)) +
  geom_point()

```























